---
title: "Bullet Data Preparation for LIME Diagnostics Paper"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document: 
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, fig.align = "center")
```

This document contains code for preparing the raw bullet data sets (training, testing, and example matching signatures) to be used the paper "Visual Diagnostics of a Model Explainer -- Tools for the Assessment of LIME Explanations".

**NOTE** The code chunks where the prepared data versions are shared have `eval = FALSE`. These code chunks must be run manually. This is done to help avoid any accidental changes of the data.

Load R packages:

```{r}
library(cowplot)
library(dplyr)
library(forcats)
library(ggplot2)
library(randomForest)
library(readr)
library(stringr)
library(tidyr)
library(zip)
```

Obtain features used when fitting the rtrees random forest:

```{r}
rf_features <- rownames(bulletxtrctr::rtrees$importance)
```

# Bullet Training Data

## Comparing Training Datasets

We are trying to determine which of two datasets were used to train `rtrees` in `bulletxtrctr`.

Load the raw version of the Hamby 172 and 252 bullet comparison data (provided by Heike and CSAFE):

```{r}
bullet_train_raw = read_csv("../data/raw/hamby-comparisons.csv.zip")
```

Load the newest training data provided by Heike (on 9/24/2020 from CSAFE data base):

```{r}
ccfs_withlands_raw = read.csv("../data/raw/CCFs_withlands.csv")
```

### Initial Comparisons

Data dimensions:

```{r}
dim(bullet_train_raw)
dim(ccfs_withlands_raw)
```

Comparing the distributions of the two datasets:

```{r}
summary(bullet_train_raw %>% filter(flag == FALSE) %>% select(rf_features))
summary(ccfs_withlands_raw %>% select(rf_features))
```

### Clean CCFs with lands

Rename some of the variables to match the structure of `bullet_train_raw`, select only variables of interest, and convert LEA labels to factors (for plotting purposes): 

```{r}
ccfs_withlands_raw_expanded <- 
  ccfs_withlands_raw %>%
  rename("study1" = "study.x", "barrel1" = "barrel.x", "bullet1" = "bullet.x", "land1" = "land.x",
         "study2" = "study.y", "barrel2" = "barrel.y", "bullet2" = "bullet.y", "land2" = "land.y") %>%
  select(
    study1, barrel1, bullet1, land1, 
    study2, barrel2, bullet2, land2,
    all_of(rf_features), 
    match
  ) %>%
  mutate(across(c(study1, barrel1, bullet1, land1, study2, barrel2, bullet2, land2), factor))
```

Create a version of the `bullet_train_raw` data with the land IDs expanded to multiple columns for plotting:

```{r}
bullet_train_raw_expanded <- 
  bullet_train_raw %>%
  separate(land_id1, c("study1", "barrel1", "bullet1", "land1")) %>%
  separate(land_id2, c("study2", "barrel2", "bullet2", "land2")) %>%
  mutate(across(where(is.character), factor)) %>%
  mutate(barrel1 = fct_relevel(barrel1, paste0("Br", 1:10)),
         barrel2 = fct_relevel(barrel2, paste0("Br", 1:10)))
```

Shows that `Hamby44` needs to be changed to `Hamby173` in `ccfs_withlands`:

```{r fig.width = 12, fig.height = 5, echo = FALSE}
plot_grid(
  bullet_train_raw_expanded %>% count(study1, study2) %>%
    ggplot(aes(
      x = study1, y = study2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) + 
    labs(title = "Hamby comparisons data"),
  ccfs_withlands_raw_expanded %>% count(study1, study2) %>%
    ggplot(aes(
      x = study1, y = study2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) +
    labs(title = "CCFs with lands data")
)
```

Shows that lettered barrels need to be changed to `BrUnk` and `Br` added to other barrels in `ccfs_withlands`:

```{r fig.width = 12, fig.height = 5, echo = FALSE}
plot_grid(
  bullet_train_raw_expanded %>% count(barrel1, barrel2) %>%
    ggplot(aes(
      x = barrel1, y = barrel2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) + 
    labs(title = "Hamby comparisons data"),
  ccfs_withlands_raw_expanded %>% count(barrel1, barrel2) %>%
    ggplot(aes(
      x = barrel1, y = barrel2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) +
    labs(title = "CCFs with lands data")
)
```

Shows that barrel letters from in `ccfs_withlands` need to be made the bullet and `B` added to all bullets:

```{r fig.width = 12, fig.height = 5, echo = FALSE}
plot_grid(
  bullet_train_raw_expanded %>% count(bullet1, bullet2) %>%
    ggplot(aes(
      x = bullet1, y = bullet2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) + 
    labs(title = "Hamby comparisons data"),
  ccfs_withlands_raw_expanded %>% count(bullet1, bullet2) %>%
    ggplot(aes(
      x = bullet1, y = bullet2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) +
    labs(title = "CCFs with lands data")
)
```

Shows that `L` needs to be added to lands in `ccfs_withlands`:

```{r fig.width = 12, fig.height = 5, echo = FALSE}
plot_grid(
  bullet_train_raw_expanded %>% count(land1, land2) %>%
    ggplot(aes(
      x = land1, y = land2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) + 
    labs(title = "Hamby comparisons data"),
  ccfs_withlands_raw_expanded %>% count(land1, land2) %>%
    ggplot(aes(
      x = land1, y = land2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) +
    labs(title = "CCFs with lands data")
)
```

Determine the letters in barrels: 

```{r}
all_barrel_labels <-
  unique(c(
    as.character(ccfs_withlands_raw_expanded$barrel1),
    as.character(ccfs_withlands_raw_expanded$barrel2)
  ))
letters <- all_barrel_labels[!(all_barrel_labels %in% 1:10)]
letters
```

Clean `ccfs_withlands` so the labels match `bullet_train_raw`:

```{r}
ccfs_withlands_clean_expanded <- ccfs_withlands_raw_expanded %>%
  mutate(
    study1 = fct_recode(study1, "Hamby173" = "Hamby44"),
    study2 = fct_recode(study2, "Hamby173" = "Hamby44"),
    bullet1 = ifelse(barrel1 %in% letters, as.character(barrel1), as.character(bullet1)),
    barrel1 = ifelse(barrel1 %in% letters, "Unk", barrel1),
    bullet2 = ifelse(barrel2 %in% letters, as.character(barrel2), as.character(bullet2)),
    barrel2 = ifelse(barrel2 %in% letters, "Unk", barrel2)
    ) %>%
  mutate(
    study1 = fct_relevel(study1, "Hamby173", "Hamby252"),
    study2 = fct_relevel(study2, "Hamby173", "Hamby252"),
    barrel1 = paste0("Br", barrel1),
    bullet1 = paste0("B", bullet1),
    land1 = paste0("L", land1),
    barrel2 = paste0("Br", barrel2),
    bullet2 = paste0("B", bullet2),
    land2 = paste0("L", land2)
  ) %>% 
  mutate(across(where(is.character), factor)) %>%
  mutate(barrel1 = fct_relevel(barrel1, paste0("Br", 1:10)),
         barrel2 = fct_relevel(barrel2, paste0("Br", 1:10)))
```

The labels agree now but we are starting to see some differences in the two datasets:

```{r fig.width = 12, fig.height = 5, echo = FALSE}
plot_grid(
  bullet_train_raw_expanded %>% count(study1, study2) %>%
    ggplot(aes(
      x = study1, y = study2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) + 
    labs(title = "Hamby comparisons data"),
  ccfs_withlands_clean_expanded %>% count(study1, study2) %>%
    ggplot(aes(
      x = study1, y = study2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) +
    labs(title = "CCFs with lands data")
)
```

The labels agree now but still can see the differences in the two datasets:

```{r fig.width = 12, fig.height = 5, echo = FALSE}
plot_grid(
  bullet_train_raw_expanded %>% count(barrel1, barrel2) %>%
    ggplot(aes(
      x = barrel1, y = barrel2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) + 
    labs(title = "Hamby comparisons data"),
  ccfs_withlands_clean_expanded %>% count(barrel1, barrel2) %>%
    ggplot(aes(
      x = barrel1, y = barrel2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) +
    labs(title = "CCFs with lands data")
)
```

The labels agree now but again, there are apparent differences between datasets:

```{r fig.width = 12, fig.height = 5, echo = FALSE}
plot_grid(
  bullet_train_raw_expanded %>% 
    count(bullet1, bullet2) %>%
    ggplot(aes(
      x = bullet1, y = bullet2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) + 
    labs(title = "Hamby comparisons data"),
  ccfs_withlands_clean_expanded %>% 
    count(bullet1, bullet2) %>%
    ggplot(aes(
      x = bullet1, y = bullet2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) +
    labs(title = "CCFs with lands data")
)
```

The labels agree now:

```{r fig.width = 12, fig.height = 5, echo = FALSE}
plot_grid(
  bullet_train_raw_expanded %>% 
    count(land1, land2) %>%
    ggplot(aes(
      x = land1, y = land2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) + 
    labs(title = "Hamby comparisons data"),
  ccfs_withlands_clean_expanded %>% 
    count(land1, land2) %>%
    ggplot(aes(
      x = land1, y = land2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) +
    labs(title = "CCFs with lands data")
)
```

Compress the land labels for `ccfs_withlands`: 

```{r}
ccfs_withlands_clean <- 
  ccfs_withlands_clean_expanded %>%
  mutate(
    land_id1 = paste(study1, barrel1, bullet1, land1, sep = "-"),
    land_id2 = paste(study2, barrel2, bullet2, land2, sep = "-")
  ) %>%
  select(land_id1, land_id2, all_of(rf_features), match)
```

### Land ID Differences

Extract the unique land IDs from the two datasets:

```{r}
ccfs_withlands_ids = unique(c(ccfs_withlands_clean$land_id1, ccfs_withlands_clean$land_id2))
bullet_train_ids = unique(c(bullet_train_raw$land_id1, bullet_train_raw$land_id1))
```

Identify the land ID in `CCFs_withlands` but not in `hamby-comparisons`:

```{r}
ccfs_withlands_ids[!(ccfs_withlands_ids %in% bullet_train_ids)]
```

Identify the land ID in `hamby-comparisons` but not in `CCFs_withlands`:

```{r}
bullet_train_ids[!(bullet_train_ids %in% ccfs_withlands_ids)]
```

## Compare Dimensions to rtrees

The dimensions of the raw data:

```{r}
bullet_train_raw %>% dim()
```

The dimension of the data once the observations with tank rash (`flag != FALSE`) are removed:

```{r}
bullet_train_raw %>% filter(flag == FALSE) %>% dim()
```
 
The number of observations in the random forest `rtrees`: 

```{r}
length(bulletxtrctr::rtrees$predicted)
```

The number of rows in the training data with the flagged observations excluded does not match the number of observations used to train `rtrees`. In the Hare, Hofmann, and Carriquiry (2017), they describe removing four land impressions that were flagged for quality assessment: 

- Barrel 6 Bullet 2-1
- Barrel 9 Bullet 2-4
- Unknown Bullet B-2
- Unknown Bullet Q-4

It is unclear which study these lands correspond to. The code below removes all comparisons that include these barrel, bullet, and land combinations:

```{r}
bullet_train_filtered <- 
  bullet_train_raw %>%
  mutate(bbl1 = str_remove(land_id1, pattern = "Hamby252-|Hamby173-"),
         bbl2 = str_remove(land_id2, pattern = "Hamby252-|Hamby173-")) %>%
  filter(!(bbl1 %in% c("Br6-B2-L1", "Br9-B2-L4", "BrUnk-BB-L2", "BrUnk-BQ-L4") | 
           bbl2 %in% c("Br6-B2-L1", "Br9-B2-L4", "BrUnk-BB-L2", "BrUnk-BQ-L4"))) %>%
  select(-bbl1, -bbl2) 
```

The dimensions now agree with the number of observations in `rtrees`:

```{r}
dim(bullet_train_filtered)
length(bulletxtrctr::rtrees$predicted) == dim(bullet_train_filtered)[1]
```

The code below removes observations based on Eric Hare's code (available [here](https://github.com/erichare/imaging-paper/blob/master/code/full_run.R)), which he thinks may be the code used to train `rtrees`:

```{r}
bullet_train_eric_filter <- 
  bullet_train_raw %>%
  separate(col = "land_id1", into = c("study1", "barrel1", "bullet1", "land1")) %>%
  separate(col = "land_id2", into = c("study2", "barrel2", "bullet2", "land2")) %>%
  filter(study1 != "Hamby173" | bullet1 != "BE") %>%
  filter(study2 != "Hamby173" | bullet2 != "BE")

# Eric's code for reference:
# filter(study.y != "Hamby44" | barrel.y != "E") %>%
# filter(study.x != "Hamby44" | barrel.x != "E") %>%
```

Here, the dimensions to not agree with the number of observations in `rtrees`:

```{r}
dim(bullet_train_eric_filter)
length(bulletxtrctr::rtrees$predicted) == dim(bullet_train_eric_filter)[1]
```

```{r}
bullet_train_eric_filter_v2 <-
  bullet_train_raw %>%
  mutate(bbl1 = str_remove(land_id1, pattern = "Hamby252-|Hamby173-"),
         bbl2 = str_remove(land_id2, pattern = "Hamby252-|Hamby173-")) %>%
  filter(!(bbl1 %in% c("Br6-B2-L1", "Br9-B2-L4", "BrUnk-BB-L2", "BrUnk-BQ-L4") | 
           bbl2 %in% c("Br6-B2-L1", "Br9-B2-L4", "BrUnk-BB-L2", "BrUnk-BQ-L4"))) %>%
  select(-bbl1, -bbl2) %>%
  separate(col = "land_id1",
           into = c("study1", "barrel1", "bullet1", "land1")) %>%
  separate(col = "land_id2",
           into = c("study2", "barrel2", "bullet2", "land2")) %>%
  filter(study1 != "Hamby173" | bullet1 != "BE") %>%
  filter(study2 != "Hamby173" | bullet2 != "BE")

dim(bullet_train_eric_filter_v2)
length(bulletxtrctr::rtrees$predicted) == dim(bullet_train_eric_filter_v2)[1]
```

Extracting the predictions on the training data from `rtrees` and using `rtrees` to make predictions on the filtered data:

```{r}
rtrees_pred = predict(bulletxtrctr::rtrees, type = "prob")[,2]

new_pred = predict(bulletxtrctr::rtrees, 
                   bullet_train_filtered %>% 
                     select(rownames(bulletxtrctr::rtrees$importance)), type = "prob")[,2]

new_pred_full = predict(bulletxtrctr::rtrees, 
                        bullet_train_raw %>%
                          select(rownames(bulletxtrctr::rtrees$importance)), type = "prob")[,2]
```

Plot comparing the two predictions (we discovered that the `predict` function from random forest returns different results when applied to only the model (out-of-bag predictions) and the training data (regular predictions), so this is not a good way to test for agreement between training data and models):

```{r}
data.frame(rtrees_pred = sort(rtrees_pred), new_pred = sort(new_pred)) %>%
  ggplot(aes(rtrees_pred, new_pred)) + 
  geom_point()
```

Plot showing distributions of predictions from `rtrees` and new predictions made on full training data (same comment here as above about not being able to use this to determine if we have the correct training data for the model):

```{r}
data.frame(data = c(rep("rtrees", length(rtrees_pred)), 
                    rep("new predictions on full data", length(new_pred_full))), pred = c(rtrees_pred, new_pred_full)) %>%
  ggplot(aes(x = pred, fill = data)) + 
  geom_density(alpha = 0.75)
```

Comparison to the data Eric Hare said was used to fit `rtrees` which has half the number of observations as the number of rtrees predictions: 

```{r}
eric <- read.csv("https://raw.githubusercontent.com/erichare/imaging-paper/master/data/data-25-25/bullet-stats.csv")
dim(eric)
```

## Preparing Data

Determine the number of total study-barrel-bullet-lands in the data:

```{r}
length(unique(bullet_train_raw %>% pull(land_id1)))
```

Determine the number of total study-barrel-bullet-lands in the data when observations with tank rash are excluded:

```{r}
length(unique(bullet_train_raw %>% filter(flag == FALSE) %>% pull(land_id1)))
```

There is only one less land id when the cases with tank rash are excluded. The code below shows that this case is `Hamby252-Br9-B2-L4` since there are no rows in the filtered data without `Hamby252-Br9-B2-L4` as one of the land ids:

```{r}
length(unique(
  bullet_train_raw %>% filter(flag == TRUE) %>% filter(
    land_id1 != "Hamby252-Br9-B2-L4",
    land_id2 != "Hamby252-Br9-B2-L4"
  ) %>% pull(land_id1)
))
```

Using the data with the flagged lands removed, select only variables of use for the paper, add a case variable, and rename `same_source` as `samesource`:

```{r}
bullet_train <- 
  bullet_train_filtered %>%
  select(-lag, -abs_lag, -overlap, -signature_length, -flag) %>%
  mutate(case = factor(1:n())) %>%
  rename("samesource" = "same_source") %>%
  select(case, land_id1, land_id2, everything())
```

Add a variable to the training data with the random forest probability in support of a match using the random forest model `rtrees` from `bulletxtrctr`:

```{r}
bullet_features <- rownames(bulletxtrctr::rtrees$importance)
bullet_train$rfscore <- 
  bulletxtrctr::rtrees %>%
  predict(bullet_train %>% select(all_of(bullet_features)), type = "prob") %>%
  data.frame() %>%
  pull(TRUE.)
```

Save the prepared data as a csv file:

```{r eval = FALSE}
write.csv(
  x = bullet_train, 
  file = "../data/bullet-train.csv", 
  row.names = FALSE
)
```

Also, save the original and prepared data as a zip files (for uploading to the GitHub repository):

```{r eval = FALSE}
zip(
  zipfile = "../data/bullet-train.csv.zip", 
  files = "../data/bullet-train.csv"
)
```

# Bullet Testing Data

Load in the raw Hamby 224 datasets (sets 1 and 11; provided by Heike and CSAFE):

```{r}
if (!file.exists("../data/raw/h224-set1-features.rds")) unzip("../data/raw/h224-set1-features.rds.zip")
if (!file.exists("../data/raw/h224-set11-features.rds")) unzip("../data/raw/h224-set11-features.rds.zip")

hamby224_set1 <- readRDS("../data/raw/h224-set1-features.rds")
hamby224_set11 <- readRDS("../data/raw/h224-set11-features.rds")
```

The bullet and land names from each of the sets are included below. This indicates that each set contains `r length(unique(hamby224_set1$bulletA))` bullets with `r length(unique(hamby224_set1$landA))` lands each.

| Set | Bullets | Lands | 
| --- | -------------- | -------------- | 
| 1 | `r unique(hamby224_set1$bulletA)` | `r unique(hamby224_set1$landA)` |
| 11 | `r unique(hamby224_set11$bulletA)` | `r unique(hamby224_set11$landA)` |

The dimensions of the datasets are:

| Set | Number of Rows | Number of Columns | 
| --- | -------------- | -------------- | 
| 1 | `r dim(hamby224_set1)[1]` | `r dim(hamby224_set1)[2]` |
| 11 | `r dim(hamby224_set11)[1]` | `r dim(hamby224_set11)[2]` |

The number of rows in each of these datasets indicates that there are some lands missing since a complete set with 3 bullets and 6 lands would contains 3 bullets x 3 bullets x 6 lands x 6 lands  = `r 6 * 6 * 3 * 3` comparisons. The figures below show that set 1 is missing land 4 from bullet Q, and set 11 is missing land2 from bullet 1 and land 4 from bullet I.

```{r fig.width = 10, fig.height = 4.5}
plot_grid(
  hamby224_set1 %>%
    count(bulletA, bulletB, landA, landB) %>%
    ggplot(aes(
      x = landA, y = landB, fill = n
    )) +
    geom_tile() +
    facet_grid(bulletB ~ bulletA) +
    scale_fill_distiller(palette = "GnBu", direction = 1) +
    theme_minimal() +
    theme(aspect.ratio = 1),
  hamby224_set11 %>%
    count(bulletA, bulletB, landA, landB) %>%
    ggplot(aes(
      x = landA, y = landB, fill = n
    )) +
    geom_tile() +
    facet_grid(bulletB ~ bulletA) +
    scale_fill_distiller(palette = "GnBu", direction = 1) +
    theme_minimal() +
    theme(aspect.ratio = 1, 
          axis.text.x = element_text(angle = 45, hjust = 1)),
  align = "hv"
)
```

Preparing the data from sets 1 and 11 for analysis:

```{r}
hamby224_set1_cleaned <-
  hamby224_set1 %>%
  # Remove unneeded variables
  select(-bullet_score,-land1,-land2,-aligned,-striae,-features) %>%
  # Rename bullet and land variables
  rename(
    bullet1 = bulletA,
    bullet2 = bulletB,
    land1 = landA,
    land2 = landB
  ) %>%
  # Create study and set variables and change levels of bullet and land
  # variables to match those in the training data
  mutate(
    study = factor("Hamby224"),
    set = factor("Set1"),
    bullet1 = recode(factor(bullet1), "1" = "B1", "2" = "B2", "Q" = "BUnk"),
    bullet2 = recode(factor(bullet2), "1" = "B1", "2" = "B2", "Q" = "BUnk"),
    land1 = paste0("L", land1),
    land2 = paste0("L", land2)
  ) %>%
  # Order the variables as desired
  select(study, set, bullet1, land1, bullet2, land2, all_of(rf_features), samesource, rfscore)

hamby224_set11_cleaned <- 
  hamby224_set11 %>%
  # Remove unneeded variables
  select(-bullet_score,-land1,-land2,-aligned,-striae,-features) %>%
  # Rename bullet and land variables
  rename(
    bullet1 = bulletA,
    bullet2 = bulletB,
    land1 = landA,
    land2 = landB
  ) %>%
  # Create study and set variables and change levels of bullet and land
  # variables to match those in the training data
  mutate(
    study = factor("Hamby224"),
    set = factor("Set11"),
    bullet1 = recode(factor(bullet1), "Bullet 1" = "B1", "Bullet 2" = "B2", "Bullet I" = "BUnk"),
    bullet2 = recode(factor(bullet2), "Bullet 1" = "B1", "Bullet 2" = "B2", "Bullet I" = "BUnk"),
    land1 = str_remove(land1, "and "),
    land2 = str_remove(land2, "and ")
  ) %>%
  # Order the variables as desired
  select(study, set, bullet1, land1, bullet2, land2, all_of(rf_features), samesource, rfscore)
```

Plots show that the number of observations is still the same after cleaning:

```{r fig.width = 10, fig.height = 4.5}
plot_grid(
  hamby224_set1_cleaned %>%
    count(bullet1, bullet2, land1, land2) %>%
    ggplot(aes(
      x = land1, y = land2, fill = n
    )) +
    geom_tile() +
    facet_grid(bullet2 ~ bullet1) +
    scale_fill_distiller(palette = "GnBu", direction = 1) +
    theme_minimal() +
    theme(aspect.ratio = 1),
  hamby224_set11_cleaned %>%
    count(bullet1, bullet2, land1, land2) %>%
    ggplot(aes(
      x = land1, y = land2, fill = n
    )) +
    geom_tile() +
    facet_grid(bullet2 ~ bullet1) +
    scale_fill_distiller(palette = "GnBu", direction = 1) +
    theme_minimal() +
    theme(aspect.ratio = 1),
  align = "hv"
)
```

Join the two cleaned Hamby 224 sets into one testing set, remove duplicate comparisons, create a case variable, and create land ids based on the study, set, bullet, and land (remove those variables afterwards):

```{r}
bullet_test_extra_vars <-
  # Join the two data sets
  bind_rows(hamby224_set1_cleaned, hamby224_set11_cleaned) %>%
  # Remove duplicate comparisons
  filter(!(bullet1 == "BUnk" & bullet2 == "B1"),
         !(bullet1 == "BUnk" & bullet2 == "B2"),
         !(bullet1 == "B2" & bullet2 == "B1")) %>%
  # Create a case and land_id variables
  mutate(
    case = factor(1:length(study)),
    land_id1 = paste(study, set, bullet1, land1, sep = "-"),
    land_id2 = paste(study, set, bullet2, land2, sep = "-")
  ) # Leave the extra variables for plotting

# Select and order variables to keep for analysis
bullet_test <- bullet_test_extra_vars %>%
  select(case, land_id1, land_id2, all_of(rf_features), samesource, rfscore)
```

Remaining comparisons included in the test data after the duplicate comparisons are removed:

```{r fig.width = 10, fig.height = 4.5}
plot_grid(
  bullet_test_extra_vars %>%
    filter(set == "Set1") %>%
    count(bullet1, land1, bullet2, land2) %>%
    ggplot(aes(
      x = land1, y = land2, fill = n
    )) +
    geom_tile() +
    facet_grid(bullet2 ~ bullet1) +
    scale_fill_distiller(palette = "GnBu", direction = 1) +
    theme_minimal() +
    theme(aspect.ratio = 1),
  bullet_test_extra_vars %>%
    filter(set == "Set11") %>%
    count(bullet1, land1, bullet2, land2) %>%
    ggplot(aes(
      x = land1, y = land2, fill = n
    )) +
    geom_tile() +
    facet_grid(bullet2 ~ bullet1) +
    scale_fill_distiller(palette = "GnBu", direction = 1) +
    theme_minimal() +
    theme(aspect.ratio = 1),
  align = "hv"
)
```

Save the test data as a csv file:

```{r eval = FALSE}
write.csv(
  x = bullet_test, 
  file = "../data/bullet-test.csv", 
  row.names = FALSE
)
```

Also, save the original and prepared data as a zip files (for uploading to the GitHub repository):

```{r eval = FALSE}
zip(
  zipfile = "../data/bullet-test.csv.zip", 
  files = "../data/bullet-test.csv"
)
```

# Example Matching Signatures

The following code trims the data provided by Heike of two matching bullet-land signatures.

Import the data (provided by Heike and CSAFE):

```{r}
if (!file.exists("../data/raw/signatures.rds")) unzip("../data/raw/signatures.rds.zip")
signatures <- readRDS("../data/raw/signatures.rds")
```

Structure of the data:

```{r}
signatures %>% str()
```

Create a land variable based on the source variable, select only variables necessary for the manuscript, and rename sig as y:

```{r}
signatures_trimmed <-
  signatures %>%
  mutate(land = c("Signature 1", "Signature 2")[as.factor(source)]) %>%
  select(land, x, sig) %>%
  rename(y = sig)
```

Export the trimmed data as a csv file:

```{r eval = FALSE}
write.csv(
  x = signatures_trimmed,
  file = "../data/example-signatures.csv",
  row.names = FALSE
)
```

Also, save the original and prepared data as a zip files (for uploading to the GitHub repository):

```{r eval = FALSE}
zip(
  zipfile = "../data/example-signatures.csv.zip", 
  files = "../data/example-signatures.csv"
)
```

